---
title: "Prediction Bound Change-Point Method"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(predictBoundCP)
```

## Overview of change-point method

This package contains the scripts necessary to find a change point by finding the point
where the observed data systematically deviate outside of, and above, the prediction bound
of a model fit to a training period before the diagnostic opportunity window.

The following steps are applied to find the change-point:

1. Specify the training period [`low`,`high`] days prior to diagnosis, where `low` 
is > the furthest point in the "feasible" opportunity window

2. Fit a model to the data over the period [`low`,`high`].

3. Generate forecasts over the period [0, `low`-1] along with a corresponding 
prediction interval

4. Find the change-point *cp* where the observed number of SSD visits exceeds the 
upper prediction bound for all points *t<cp* prior to diagnosis

## Example

### SSD Visit Count Data

The package contains a count dataset corresponding to SSD visits prior to HSV encephalitis. 
The change-point functions require a count dataset with two variables: `days_since_dx` 
is the number of days since diagnosis (negative values before diagnosis) and `n`
is the number of visits

```{r}
visit_counts
```
The following figure depicts the trend in visit counts prior to HSV diagnosis, for
the 180 days prior to diagnosis
``` {r}
plot_ssd_curve(count_data = visit_counts,
               max_window = 180)
```

### Finding a change-point

The main function is `fit_cp_model()`. This function takes an SSD visit count dataset 
(*visit_count*), *low* and *high* values for the training window, a *model* to fit 
over the training window, the confidence *level* for the prediction bound. The 
function also has an option to return a plot of the fitted model along with prediction
bounds and the change-point.

The current options for models includes:

* `lm` - Linear Model
* `exp` - Exponential Model
* `lm_period` - Linear Model with 7-day Periodicity
* `exp_period` - Exponential Model with 7-day Periodicity
* `auto_arima` - Auto ARIMA Model

``` {r warning=FALSE, message=FALSE}
# Linear Model
mod1 <- fit_cp_model(count_data = visit_counts,
                     low = 40, high = 180,      # training window
                     model = "lm",              # model
                     level = .95,               # prediction bound
                     plot = TRUE)

# Exponential Model
mod2 <- fit_cp_model(count_data = visit_counts,
                     low = 40, high = 180,          # training window
                     model = "exp",                 # model
                     level = .95,                   # prediction bound
                     plot = TRUE)

# Linear Model with Periodicity
mod3 <- fit_cp_model(count_data = visit_counts,
                     low = 40, high = 180,          # training window
                     model = "lm_period",           # model
                     level = .95,                   # prediction bound
                     plot = TRUE)

# Exponential Model with Periodicity
mod4 <- fit_cp_model(count_data = visit_counts,
                     low = 40, high = 180,          # training window
                     model = "exp_period",          # model
                     level = .95,                   # prediction bound
                     plot = TRUE)

# Auto ARIMA Model
mod5 <- fit_cp_model(count_data = visit_counts,
                     low = 40, high = 180,          # training window
                     model = "auto_arima",          # model
                     level = .95,                   # prediction bound
                     plot = TRUE)
```

We can compare the results of the different models

``` {r}
gridExtra::grid.arrange(mod1$plot, mod2$plot,
                        mod3$plot, mod4$plot, 
                        nrow = 2)

mod5$plot
```

### Profiling different change-points

The function `fit_cp_profile()` creates a profile of change-points using a range
of different models, training windows and prediction bounds.

``` {r warning=FALSE, message=FALSE}
profile_res <- fit_cp_profile(count_data = visit_counts,
                              low = 40,
                              high = 180)
```

We can view the change-points and the results across the various parameters:
``` {r}
profile_res
```

Or we can plot the results across different lower bounds used in the training data:
``` {r}
profile_res %>% 
  ggplot2::ggplot(aes(low,cp,color=model)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~bound) +
  ggplot2::theme_minimal() + 
  ggplot2::theme(legend.position="bottom")
````
